A Discrepancy-Based Algorithm for Extended Inversion

### 2020-08-03 repro/papers/discrep/paper.tex
interim

@@

Note that for $w \in \lW$, $m \in M$,
$\mbox{supp }F[m]w \subset [ m_{\rm min}r-\lambda,
m_{\rm max}r+\lambda] \cap
[t_{\rm min},t_{\rm max}]$. Also, $\lF[m]$ is coercive for all $m \in
M$ (so that $w$ is uniquely determined) only if
$[ m_{\rm min}r-\lambda,m_{\rm max}r+\lambda] \subset
[t_{\rm min},t_{\rm max}]$. Define $\lambda_{\rm max}
>0$, so that $M, r,$ and $T$ by
\begin{equation}
  \label{eqn:fullrec}
  [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
  \subset [t_{\rm min},t_{\rm max}].
\end{equation}
Then for 
$m \in M$, $0 <\lambda \le \lambda_{\rm max}$, $F[m] \in {\cal B}(\lW,D)$ is coercive.

@@

\begin{theorem}
  \label{thm:epjnonoise}
  Assume the hypotheses of Theorem \ref{thm:norminv}, item 3. Then 
\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= 8 \pi^2 r^2 \alpha^4\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m^*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,  a(t-(m-m_*)r)^2 \left(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2\right)^{-2}w_*(t)^2.
\end{equation}
so finally
\begin{equation}
\label{eqn:expjgen}
\tJa[m;d] = =\frac{\alpha^2}{2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m^*)r)^2)^{-1}w_*(t)^2.
\end{equation}
\end{theorem}

I willl now examine two possible choices of penalty multipler $a$ in
defintion \ref{eqn:annmult}.

@@

A second possible 
penalty operator penalizes energy away from 
$t=0$: choose $\tau > 0$ and set 
\begin{equation}
  \label{eqn:ann}
  a(t) = \min(|t|, \tau). 
\end{equation}
The cutoff $\tau$ will be chosen large enough to be effectively inactive: 
specifically, hindsight suggests 
\begin{equation}
  \label{eqn:taudef}
  \tau = \max\{|t_{\rm min}-m_{\rm min}r|,|t_{\rm min}-m_{\rm max}r|, |t_{\rm max}-m_{\rm min}r|, |t_{\rm max}-m_{\rm max}r|\}. 
\end{equation}
This 
particular annihilator has been employed in earlier papers on extended 
source inversion 
\cite[]{Plessix:00a,LuoSava:11,Warner:14,HuangSymes:SEG15a,Warner:16,HuangSymes:GEO17}. It
satisfies condition \ref{eqn:abnd} with $C=\lambda$.

@@

\begin{theorem}
  \label{thm:rampgood}
  For $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
  with $\tau$ given by equation \ref{eqn:taudef}, all
  stationary points of $\tJa$ lie in the interval
  \[
    [m_*-\lambda/r, m_*+\lambda/r].
  \]
\end{theorem}
\begin{proof}
  As observed before, $\mbox{supp }\aw[m;d] \subset [t_{\rm
    min}-mr,t_{\rm max}-mr]\subset [-\tau,\tau]$, with $\tau$ defined
  in \ref{eqn:taudef}. Therefore for $t \in  [t_{\rm min}-mr,t_{\rm
    max}-mr]$, $a(t) = |t|$, and reason analogous to the proof of
  Theorem \ref{thm:norminv}, item 2, shows that
  \begin{equation}
    \label{eqn:solnnon}
    \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 t^2\right)^{-1}w_*\left(t+(m-m_*)r\right) 
  \end{equation}
  Straightforward modification of the proof of Theorem
  \ref{thm:epjnonoise} gives
  \begin{equation}
    \label{eqn:expj}
    \tJa[m;d] =\frac{\alpha^2}{2}\int\,dt\,(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
    (t-(m-m^*)r)^2)^{-1}w_*(t)^2.
  \end{equation}
  A brief calculation yields
  \begin{equation}
    \label{eqn:gradfinal}
    \frac{\partial}{\partial m}\tJa[m;d] = -r\alpha^2\int \,dt\,\frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}w_*(t+(m-m_*)r)^2.
  \end{equation}
  Recall that $w_*(t+(m-m_*)r)$
  vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
  RHS of equation \ref{eqn:gradfinal} can be re-written
  \[
    = -r\alpha^2\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
    \,dt\, \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}w_*\left(t+(m-m_*)r\right)^2
  \]
  If $m > m_*+\lambda/r$, then the entire interval of integration is a proper
  subset of the negative half-axis. Consequently the first factor in the
  integrand satisfies
  \[
    \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}< 0
  \]
  over the interval of integration. The integral is
  negative, so the derivative
  is positive. Similar reasoning applies to the case $m <
  m_*-\lambda/r$. Thus
  \begin{eqnarray}
    m > m_*+\lambda/r & \Rightarrow & \frac{\partial \tJa}{\partial
                                      m}[m;d]> 0; \nonumber\\
    m < m_*-\lambda/r & \Rightarrow & \frac{\partial \tJa}{\partial
                                      m}[m;d]< 0.
                                      \label{eqn:leftright}
  \end{eqnarray}
\end{proof}

@@

\subsection{The Effect of Noise}
Suppose that the data trace $d$ takes the form
\begin{equation}
  \label{eqn:defdatanoisy}
  d = F[m_*]w_* + n
\end{equation}
with $m_* \in M, w_* \in \lW$ as before, and noise trace $n \in
D$. Since no support assumptions can be made about $n$, equation
\ref{eqn:normsol} implies that $\aw[m;d] \notin \lW$ for any value of
$\alpha$.  Therefore minimization of $\tJa$ cannot by itself yield a
solution of the inverse problem as defined in the problem statement
\ref{eqn:probstat0}.

>>>>

Note also that $F[m]$ is surjective for every $m \in M$.

In computational practice, $W$ will have to be replaced by a
finite-dimensional subspace of $L^2(\bR)$. Many such choices will
implicitly limit the support of $w \in W$ to a bounded interval, say
$[T_{\rm min},T_{\rm max}]$. To maintain the surjective property,
these bounds should be chosen so that $[t_{\rm min}, t_{\rm max}]
\subset [T_{\rm min}+mr,T_{\rm max}+mr]$ for all $m \in M$, that is,
\begin{eqnarray}
  \label{eqn:dombounds}
  t_{\rm min} &\ge & T_{\rm min}+m_{\rm max}r,\nonumber\\
  t_{\rm max} &\le & T_{\rm max}+m_{\rm min}r.
\end{eqnarray}
We will ignore these computational necessities in this work,
maintaining the defintion $W=L^2(\bR)$.

@@

Note that for $w \in \lW$, $m \in M$,
$\mbox{supp }F[m]w \subset [ m_{\rm min}r-\lambda,
m_{\rm max}r+\lambda] \cap
[t_{\rm min},t_{\rm max}]$. Also, $\lF[m]$ is coercive for all $m \in
M$ (so that $w$ is uniquely determined) only if
$[ m_{\rm min}r-\lambda,m_{\rm max}r+\lambda] \subset
[t_{\rm min},t_{\rm max}]$. Define $\lambda_{\rm max}
>0$, so that $M, r,$ and $T$ by
\begin{equation}
  \label{eqn:fullrec}
  [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
  \subset [t_{\rm min},t_{\rm max}].
\end{equation}
Then for 
$m \in M$, $0 <\lambda \le \lambda_{\rm max}$, $\lF[m] \in {\cal B}(\lW,D)$ is coercive.

@@

\begin{theorem}
  \label{thm:epjgen}
  Assume the hypotheses of Theorem \ref{thm:norminv}. Then
  \begin{equation}
  \label{eqn:residnormgen}
  e[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,(4\pi r \alpha a(t-mr))^4(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)^2
\end{equation}
\begin{equation}
  \label{eqn:anninormgen}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,(4\pi r a(t-mr))^2(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)
\end{equation}

\begin{equation}
  \label{eqn:expjgen}
\tJa[m;d] = \frac{1}{2}\int\,dt\,(4\pi r \alpha a(t-mr))^2(1+(4\pi r \alpha 
a(t-mr))^2)^{-1}d(t)^2. 
\end{equation}
Finally, if $a \in W^{1,1}(\bR)$ (distributions with integrable 
derivatives), then $\tJa[\cdot;d]$ is differentiable, and 
\begin{equation}
  \label{eqn:dexpjgen}
  \frac{d}{dm}\tJa[m;d] = -(4 \pi r \alpha)^2 \int \,dt \, 
  \left(a\frac{da}{dt}\right)(t-mr)(1+(4\pi r \alpha 
  a(t-mr))^2)^{-2}d(t)^2. 
\end{equation}
\end{theorem}

\begin{corollary}
  \label{thm:epjnonoise}
  Assume the hypotheses of Theorem \ref{thm:norminv}, item 3. Then 
\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= 8 \pi^2 r^2 \alpha^4\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m^*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,  a(t-(m-m_*)r)^2 \left(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2\right)^{-2}w_*(t)^2.
\end{equation}
so 
\begin{equation}
\label{eqn:expjnonoise}
\tJa[m;d] = \frac{\alpha^2}{2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{-1}w_*(t)^2. 
\end{equation}
Finally, if $a \in W^{1,1}(\bR)$, then $\tJa[\cdot;d]$ is differentiable, and 
\begin{equation}
  \label{eqn:dexpjnonoise}
  \frac{d}{dm}\tJa[m;d] = -r \alpha^2 \int \,dt \, 
  \left(a\frac{da}{dt}\right)(t-(m-m_*)r)(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{-2}w_*(t)^2. 
\end{equation}
\end{corollary}

I willl now examine two possible choices of penalty multipler $a$ in
definition \ref{eqn:annmult}.

@@

A second possible 
penalty operator penalizes energy away from 
$t=0$: choose $\tau > 0$ and set 
\begin{equation}
  \label{eqn:ann}
  a(t) = \min(|t|, \tau). 
\end{equation}
The cutoff $\tau$ will be chosen large enough to be effectively inactive: 
specifically, hindsight suggests 
\begin{equation}
  \label{eqn:taudef}
  \tau = \max\{|t_{\rm min}-m_{\rm min}r|,|t_{\rm min}-m_{\rm max}r|, |t_{\rm max}-m_{\rm min}r|, |t_{\rm max}-m_{\rm max}r|\}. 
\end{equation}
This 
particular annihilator has been employed in earlier papers on extended 
source inversion 
\cite[]{Plessix:00a,LuoSava:11,Warner:14,HuangSymes:SEG15a,Warner:16,HuangSymes:GEO17}. It
satisfies condition \ref{eqn:abnd} with $C=\lambda$, and is a member
of $W^{1,1}(\bR)$, so that the resulting objective $\tJa$ is
differentiable, with derivative given by equation \ref{eqn:dexpgen},
or \ref{eqn:dexpjnonoise} in the noise-free case.

@@

\begin{theorem}
  \label{thm:rampgood}
  For $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
  with $\tau$ given by equation \ref{eqn:taudef}, all
  stationary points of $\tJa$ lie in the interval
  \[
    [m_*-\lambda/r, m_*+\lambda/r].
  \]
\end{theorem}
\begin{proof}
  As observed before, $\mbox{supp }\aw[m;d] \subset [t_{\rm
    min}-mr,t_{\rm max}-mr]\subset [-\tau,\tau]$, with $\tau$ defined
  in \ref{eqn:taudef}. Therefore, $a(t) = |t|$ $a a'(t) = t$ in the
  support of the integrand on the RHS of equation
  \ref{eqn:depjnonoise}, which therefore 
  becomes (after change of integration variable)
  \begin{equation}
    \label{eqn:gradfinal}
    \frac{d}{dm}\tJa[m;d] = -r \alpha^2 \int \,dt \, 
  t(1+(4\pi r)^2 \alpha^2 
  t^2)^{-2}w_*(t+(m-m_*))^2.
  \end{equation}
  Recall that $w_*(t+(m-m_*)r)$
  vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
  RHS of equation \ref{eqn:gradfinal} can be re-written
  \[
    = -r\alpha^2\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
    \,dt\, \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}w_*\left(t+(m-m_*)r\right)^2
  \]
  Suppose that $\mu < \lambda$ and $w_* \in W_{\mu}$. Then for
  $t+(m-m_*)r \in \mbox{supp }w_*$, $t\le\mu-(m-m_*)r$.
  If $m > m_*+\lambda/r$, then $t >\mu-\lambda<0$, so the support of
  the $[-\mu-(m-m_*)r, \mu-(m-m_*)r]$ is a proper
  subset of the negative half-axis. Consequently the first factor in the
  integrand satisfies
  \[
    \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}< 0
  \]
  over the interval of integration. The integral is
  negative, so the derivative
  is positive. In fact, the quantity displayed above is
  \[
    \le 
  \]
  Similar reasoning applies to the case $m <
  m_*-\lambda/r$. Thus
  \begin{eqnarray}
    m > m_*+\lambda/r & \Rightarrow & \frac{\partial \tJa}{\partial
                                      m}[m;d]> 0; \nonumber\\
    m < m_*-\lambda/r & \Rightarrow & \frac{\partial \tJa}{\partial
                                      m}[m;d]< 0.
                                      \label{eqn:leftright}
  \end{eqnarray}
\end{proof}

@@

\subsection{The Effect of Noise}
Suppose that the data trace $d$ takes the form
\begin{equation}
  \label{eqn:defdatanoisy}
  d = F[m_*]w_* + n = d_*+n,
\end{equation}
with $m_* \in M, w_* \in \lW$ as before, and noise trace $n \in
D$. Since no support assumptions can be made about $n$, equation
\ref{eqn:normsol} implies that $\aw[m;d] \notin \lW$ for any value of
$\alpha$.  Therefore minimization of $\tJa$ cannot by itself yield a
solution of the inverse problem as defined in the problem statement
\ref{eqn:probstat0}.

Recall that $F[m]$ is surjective for every $m \in M$. Define for each
$m \in M$
\begin{equation}
  \label{eqn:defnoisewavelet}
  w_{n}[m] = \mbox{arg}\min_w\{\|w\|: F[m]w=n\}.
\end{equation}
That is,
\[
  w_n[m](t) =
  \left\{
    \begin{array}{c}
      4\pi r n(t+mr),\,t \in [t_{\rm min}-mr,t_{\rm max}-mr];\\
      0, \mbox{ else.}
    \end{array}
  \right.
\]
Set
\begin{equation}
  \label{eqn:defwflat}
  w_{\flat} = w_* + w_n[m_*];
\end{equation}
then
\begin{equation}
  \label{eqn:datacomplete}
  d = F[m_*]w_{\flat}.
\end{equation}
That is, we can treat the data as if it were noise-free, for the price
of altering the wavelet. The altered wavelet is of course not a member
of $\lW$.

From the proof of Theorem \ref{thm:norminv}, for the choice
$a(t)=\min(|t|,\tau)$ with $\tau$ defined in \ref{eqn:taudef},
\begin{equation}
  \label{eqn:solnoise}
  \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 t^2\right)^{-1}w_{\flat}\left(t+(m-m_*)r\right).
\end{equation}
and from Theorem \ref{thm:epjnonoise}
\begin{equation}
\label{eqn:expjgen}
\tJa[m;d] = =\frac{\alpha^2}{2}\int\,dt\,(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  (t-(m-m^*)r)^2)^{-1}w_{\flat}(t)^2.
\end{equation}
whence via the same short computation as before
\[
  \frac{\partial}{\partial m}\tJa[m;d] = -r\alpha^2\int \,dt\,\frac{t-(m-m_*)r}{(1+(4\pi r)^2\alpha^2 (t-(m-m_*)r)^2)^2}w_{\flat}(t)^2. 
\]
\[
  = -r\alpha^2 \int \, dt \, (t-(m-m_*)r)\aw[m;d](t)^2 
\]  
\[
  = -r\alpha^2\left\{ \int_{-\lambda}^{\lambda} \, dt \, 
    (t-(m-m_*)r)(w_*(t)^2 +2w_*(t)w_n[m_*](t))\right\}
\]
\begin{equation}
  \label{eqn:gradnoise}
-r\alpha^2\int_{-\infty}^{\infty} \, dt \, a(t-(m-m_*)r) w_n[m_*](t)^2.
\end{equation}
The last line recalls the definition of $a$ for this example
(\ref{eqn:deftau}).

We will follow the pattern laid down in the proof of Theorem
\ref{thm:rampgood}. However to obtain an estimate depending on some
measure of noise, it is necessary to explicitly quantify the bounds
within which stationary points of $\tJa$ are located. Pick $\dlam >0$,
and suppose that $|m-m_*|r > \lambda + \dlam$. Then for $t \in
[-\lambda,\lambda]$,
\begin{eqnarray*}
  (m-m_*)r >   \lambda +\dlam & \Rightarrow & t-(m-m_*)r < -\dl, \\
  (m-m_*)r < -\lambda +\dlam & \Rightarrow & t-(m-m_*)r >   \dl.
\end{eqnarray*}
In the first case,
